Family Income and Educational Returns in China, 2010-2020: Higher Returns Favor the Wealthy, but Gaps Hold Steady

### 2024-12-03 paper/paper.qmd
updated paper and references

---
title: "A Logistic Regression Prediction of 2024 Election Odds Gives VP Harris a 38.3% Winrate"
author: 
  - Andrew Goh
  - Yisu Hou
thanks: "Code and data are available at: [https://github.com/YisuHou1/US_Election_Statistics](https://github.com/YisuHou1/US_Election_Statistics)."
date: today
date-format: long
abstract: "The 2024 Presidential Election, run between the Democratic candidate Vice President Kamala Harris and Republican candidate Former President Donald Trump, is both widely predicted to be a close race [@538data] and considered to be the “most important election of your life” by a majority of likely voters this election season, as estimated by Quinnipiac [@quinnipiac]. To predict the results of this election, we use a FiveThirtyEight poll-of-polls dataset [@538data], fitting multinomial logistic regression models with the data both at a national level and for each of the 7 swing states. Using these estimates for the probability of Harris winning the popular votes in each of those categories, we find that Vice President Harris has a 38.3% probability of winning the election in the context of the electoral college, a statistically significant result after considering margin of error and various biases."
format: pdf
header-includes:
  - \usepackage{float}
  - \usepackage{placeins}
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(caret)
library(glmnet)
library(nnet)
library(lubridate)
library(pROC)
library(stringr)
library(gt)
library(broom)
library(knitr)
```

@@

With most polls maintaining close (<1) odds at both the national and statewide level for the seven swing states [@538data; @nyt2024polls], the 2024 Presidential Elections are proving to be an extremely volatile one, with poll predictions regularly flipping and disagreeing with each other, even up until the day before Election Day. The volatility even extends beyond the polling itself – since the January 6 Capitol Attack, prominent figures supporting both parties have painted the 2024 elections specifically as one that underpins the very democratic integrity of the country. Elon Musk, for example, has openly expressed concerns of a Harris win leading to a one-party country [@musk2024x], and Zack Beauchamp from Vox has described a Trump re-election as an “extinction-level threat to democracy” [@vox]. In short, the stakes are high, at least as perceived in the eyes of voters [@quinnipiac].

In order to arrive at such a conclusion, we used poll-by-polls data from FiveThirtyEight [@538data] to fit multinomial logistic regression models for both the nation as a whole and each of the seven swing states. Taking into consideration the mispredictions by pollsters for both the 2016 and 2020 elections, where Trump won the presidential race against Clinton in a 304-227 landslide despite predictions universally predicting a landslide win in the opposite direction [@weighting2], we regress on the characteristics of the poll/pollster such as transparency rating and sponsors and correct the results for systematic polling biases based on past election data, an approach predicted to yield better results this year [@weighting].

Our estimand, or what we seek to predict, is the probability that Kamala Harris wins the election on the 6th of November.

After obtaining the probabilities that Harris wins each swing state via randomised sampling using our fitted models, we then analysed these probabilities in the context of the electoral college, i.e. the probability that Harris can obtain 44 votes from the swing states, which would all but guarantee her step into office [@nyt2024polls]. We find that Harris has a 38.3% chance of winning the general election, suggesting a likely Trump advantage after accounting for the margin of error in all related processes and observations.

The body of the paper contains the data, model, results, and discussion sections. In order, @sec-data contains an overview of the data used, the variables involved and the tools used in our analysis and interpretation of the data, @sec-model describes the model in detail and other related miscellaneous musings, @sec-results contains a run-down and explanation of the results and findings from the model, and @sec-discussion contains a discussion of the implications and limitations of the result, among others. The appendices contain a sample of an idealised methodology and survey, as well as a case study of the sample and methodology of Quinnipiac’s Pennsylvania poll [@quinnipiac; @quinnipiacmethodology].

@@

We use the statistical programming language R [@R] for the graphing, analysis and presentation of the project as a whole. Caret [@caret], glmnet [@glmnet; @glmnet2; @glmnet3], nnet [@nnet], lubridate [@lubridate] and pROC [@pROC] were used directly in the analysis of the data. Tidyverse [@tidyverse], stringr [@stringr], dplyr [@dplyr], gt [@gt], knitr [@knitr], and styler [@styler] were used in the presentation and/or styling of the data, graphs, and paper as a whole.

In order to build a model that predicts the results of the 2024 election, we use survey/poll data on likely voters to inform our insights regarding likely election results, an approach widely accepted in political contexts [@politicalpolling]. We use a poll-of-polls dataset from 538 [@538data], consisting of aggregated polling data from FiveThirtyEight's 2024 presidential election polling database from various national and state-level polls from pollsters all around the country, recording information such as the date when the poll concluded, the entity conducting the poll, and the estimated support rate of the candidate as given by the poll. The data spans national and state-level polls, with particular focus on key battleground states that could determine the Electoral College outcome. This dataset was chosen over alternatives like RealClearPolitics because of 538's comprehensive methodological adjustments and transparent quality standards for included polls.

@@

The "from-voter-to-data" process for the 538 dataset involves three processes: measurement, where voter responses are collected and adjusted to best provide insight into a true support percentage population parameter, data collection, where data from polls is obtained and cleaned by 538 to only include polls adhering to their set of standards, and aggregation, where additional scores/variables are given to each poll to reflect their accuracy, transparency and bias levels with regards to polling scores [@5381].

The measurement process involves several sequential steps, beginning with initial data collection where pollsters conduct surveys using various modes such as phone, online, or mixed-mode approaches to gather voter preferences. Sample processing follows, involving likely voter screens to identify probable voters, demographic weighting to match population benchmarks, and various adjustments to ensure representativeness [@quinnipiac; @newyorktimes].

The data collection process involves obtaining data from various election polls from around the country, given that they meet specific methodological criteria. Each included poll must provide clear documentation of pollster identity, survey dates, and sampled population, maintain a minimum sample size of 100 respondents, and demonstrate transparent methodology including polling mode, sample source, and weighting procedures. Scientific sampling methods attempting to achieve representative samples are mandatory, as is the disclosure of poll sponsorship and funding sources. The dataset explicitly excludes non-scientific polls lacking representative sampling, MRP-smoothed data, recontact surveys, DIY polls from nonprofessional sources, polls with leading questions or hypothetical matchups, and subsamples from multi-state polls without geographic verification [@5382].

The aggregation phase, finally, involves rating polls by pollster polling quality (ranging from 0.5 to 3.0 stars), with polls with higher historical accuracy, lower consistent partisan biases, and higher transparency of methodology boasting higher ratings [@5383]. Transparency is also a rating and is done through a 10-question yes/no checklist on how much of the methodology is available to the public - note how this is completely separate from the quality of the methodology itself. Each question is graded on a 0/0.5/1 scale, for a combined transparency score that ranges from 0 to 10. Most polls in our dataset miss the mark on one or two criteron, netting a final transparency score of 9 - this is partially due to the data cleaning process removing many relatively worsely conducted polls.

Several important measurement limitations to FiveThirtyEight's methodology must be acknowledged. These include temporal gaps between polls creating discontinuous measurement of voter sentiment, response rates and participation biases potentially skewing samples, and varying geographic coverage with swing states being overrepresented. Additionally, third-party candidate treatment varies across polls, and is difficult to standardise (with respect to other data points) consistently.

@@

The outcome variable is the level of support for the two leading US presidential candidates, Vice President Kamala Harris and Former President Donald Trump, recorded as a percentage of likely voters. For the national-level analysis, the support percentage represents the percentage of polled individuals who responded in favor of a specific candidate to a question that asked about their preferred presidential candidate in a national-level poll. For the regional datasets, the support percentage represents the percentage of likely voters who preferred a specific candidate in a poll that targets a specific state or congressional district of the United States. @fig-outcome shows the distribution of support percentages in the aggregated polling data from 2020 and 2024 presidential elections.

@@

trump_2024 <- read_csv(here::here("data/02-analysis_data/trump_2024_analysis_data.csv"))
harris_2024 <- read_csv(here::here("data/02-analysis_data/harris_2024_analysis_data.csv"))
trump_2024_lower <-
  read_csv(here::here("data/02-analysis_data/trump_2024_analysis_data_lower.csv"))
harris_2024_lower <-
  read_csv(here::here("data/02-analysis_data/harris_2024_analysis_data_lower.csv"))
trump_2020 <- read_csv(here::here("data/02-analysis_data/trump_2020_analysis_data.csv"))
biden_2020 <- read_csv(here::here("data/02-analysis_data/biden_2020_analysis_data.csv"))

# aggregated dataset for summary statistics
# datasets with lower numeric scores already includes polls with a score of 3.0
all_combined <- bind_rows(
  harris_2024_lower, trump_2024_lower,
  biden_2020, trump_2020
)
```

@@

ggplot(all_combined, aes(x = pct)) +
  geom_histogram(
    binwidth = 1, # Adjust bin width for granularity
    fill = "#2c7bb6", # Aesthetic fill color (steel blue)
    color = "white", # Border color for bins
    alpha = 0.8 # Transparency of the fill
  ) +
  labs(
    x = "Percentage Support",
    y = "Number of Polls"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  ) +
  # Add a vertical line for the mean percentage support
  geom_vline(
    aes(xintercept = mean(pct, na.rm = TRUE)),
    color = "#d73027", # Aesthetic color (red)
    linetype = "dashed",
    size = 1
  ) +
  # Annotate the mean value on the plot
  annotate(
    "text",
    x = mean(all_combined$pct, na.rm = TRUE) - 7, # Position text slightly to the right of the mean line
    y = Inf, # Position text at the top of the plot
    label = paste("Mean:", round(mean(all_combined$pct, na.rm = TRUE), 1), "%"),
    vjust = 2, # Vertical adjustment
    color = "#d73027",
    size = 5,
    fontface = "bold"
  )
```

@@

### End Date
```{r}
#| label: fig-dates-past
#| fig-cap: Distribution of Poll End Date, 2020 Cycle
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined %>% filter(cycle == 2020), aes(x = end_date)) +
  geom_histogram(binwidth = 7, fill = "#1f78b4", color = "white", alpha = 0.7) +
  labs(
    x = "End Date",
    y = "Number of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
\vspace{1em}
\FloatBarrier

```{r}
#| label: fig-dates-current
#| fig-cap: Distribution of Poll End Date, 2024 Cycle
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined %>% filter(cycle == 2024), aes(x = end_date)) +
  geom_histogram(binwidth = 7, fill = "#1f78b4", color = "white", alpha = 0.7) +
  labs(
    x = "End Date",
    y = "Number of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
\FloatBarrier
The end date indicates the date that a poll ends on. The dates are graphed separately between the 2020 and 2024 cycles for ease of viewing. Notice how there is a disproportionately large uptick in the number of polls that end on the end of a month.

@@

Sample size, finally, is a measure of the number of participants in each poll. Around 1000 participants seems to overwhelmingly be the final survey size that pollsters decide on, with only single-digit numbers of polls (with rating >= 2.5) boasting more than four thousand respondents.

@@

### Logistic models
Define $y_i$ as the binary outcome variable indicating whether Harris wins (1) or not (0) for the $i^{th}$ observation. $x_{1i}$ is the end_date for the $i^{th}$ observation. $x_{2i}$ is the binary value of has_sponsor for the $i^{th}$ observation. $x_{3i}$ is the transparency_score for the $i^{th}$ observation. $x_{4i}$ is the sample_size for the $i^{th}$ observation.

@@

\begin{align} y_i &\sim \text{Bernoulli}(p_i) 
\\ \text{logit}(p_i) &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i} \end{align}

Line (1) specifies that the outcome variable $y_i$ follows a Bernoulli distribution with probability $p_i$ of success (i.e., Harris winning). Line (2) links the linear combination of predictors to the probability $p_i$, where $\beta_0$ is the intercept term, and $\beta_1, \beta_2, \beta_3, \beta_4$ are coefficients corresponding to each predictor variable above.

@@

### Multiple linear regression models
Define $y_i$ as the percentage national support for Harris or Trump for the $i^{th}$ observation. $x_{1i}$ is the end_date for the $i^{th}$ observation. $x_{2i}$ is the binary value of has_sponsor for the $i^{th}$ observation. $x_{3i}$ is the transparency_score for the $i^{th}$ observation. $x_{4i}$ is the sample_size for the $i^{th}$ observation.

@@

\begin{align} y_i &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i} \end{align}

Where $\beta_0$ is the intercept term, and $\beta_1, \beta_2, \beta_3, \beta_4$ are coefficients corresponding to each predictor variable.

@@

roc_obj_adj <- roc(
  combined_national_adj$harris_win,
  combined_national_adj$harris_win_prob
)
auc_value_adj <- auc(roc_obj_adj)
```
For the analysis of the national popular vote, the logistic model of adjusted polls possess an AUC value of 0.98, and the logistic model of unadjusted polls have an AUC of 0.94, indicating high reliability of categorization. Their respective ROC plots are presented as @fig-log and @fig-logadj.

@@

mean_probabilities %>%
  gt() %>%
  fmt_number(
    columns = vars(Harris_Win_Probability),
    decimals = 2
  ) %>%
  cols_label(
    State = "State",
    Harris_Win_Probability = "Harris Win Probability (%)",
    electoral_votes = "Electoral Votes"
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 12,
    table.border.top.width = px(2),
    table.border.bottom.width = px(2)
  )
```
Combining with the fact that Harris requires 44 electoral votes to win and Trump requires 51, we have found using the probability function that Harris has a 38.3 chance to acquire more than 44 electoral votes, according to the probabilities and electoral votes in @tbl-result. On the other hand, Trump has a 61.7 probability to acquire more than 51 votes. Thus, our conclusion is that while both candidates possess a significant chance of winning, Trump has the greater probability of acquiring the necessary electoral votes from the seven swing states on election day.

@@

# create the plot that contains scattered points of adjusted percentages
# and lines for fitted values
ggplot(combined_data_adj, aes(x = end_date, y = pct_adj, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
\FloatBarrier

@@

```{r}
#| label: fig-lr
#| fig-cap: Unadjusted Popular Vote Prediction, 2024
#| echo: false
#| message: false
#| warning: false
# Unadjusted support percentages
ggplot(combined_data, aes(x = end_date, y = pct, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
\FloatBarrier

@@

# Discussion {#sec-discussion   }

@@

# Appendix {-}
## Pollster methodology overview and evaluation
The Quinnipiac University Poll conducts independent polling in swing states. To analyze their methodology, we look specifically into their October 2024 Pennsylvania polls [@quinnipiac].

@@

## Idealised methodology
With a $100,000 budget, our approach focuses on producing accurate state-level estimates in key battleground states, which would then inform our national forecast. We prioritize Pennsylvania, Michigan, Wisconsin, Georgia, Arizona, and Nevada, allocating resources proportionally based on each state's electoral importance and expected margin of victory.

Our sampling strategy employs both probability and non-probability methods. Probability sampling (60% of budget, i.e. $60k) means every member of our target population has a known, non-zero chance of being selected - this allows us to calculate proper margins of error and make statistical inferences about the population. For this, we use dual-frame random digit dialing (RDD) for phone surveys and address-based sampling (ABS) for mail-to-web recruitment. RDD involves generating random phone numbers within active area codes, while ABS uses the U.S. Postal Service's delivery database as a sampling frame, both proven methods of sampling [@cummings]. The ABS approach helps reach households without reliable phone access. We stratify our sample by geography, demographics, and previous voting patterns to ensure representation across key subgroups - meaning we divide the population into these subgroups and sample from each independently.

>>>>

---
title: "Family Income and Educational Returns in China, 2010-2020: Higher Returns Favor the Wealthy, but Gaps Hold Steady"
author: 
  - Yisu Hou
thanks: "Code and data are available at: [https://github.com/YisuHou1/Educational_Returns_Analysis](https://github.com/YisuHou1/Educational_Returns_Analysis)."
date: today
date-format: long
abstract: "Economic returns to education are a key driver of socioeconomic mobility, yet they are not distributed equally across all income groups. Using data from the China Family Panel Studies (CFPS) between 2010 and 2020, this paper examines how family income influences the financial benefits individuals gain from education. The analysis reveals that individuals from wealthier families consistently enjoy significantly higher income returns from education compared to their lower-income counterparts, though this gap did not widen over the decade studied. These findings highlight persistent inequality in education's economic impact, emphasizing the need for policies that ensure education serves as a more effective equalizer in society."
format: pdf
header-includes:
  - \usepackage{float}
  - \usepackage{placeins}
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(caret)
library(glmnet)
library(nnet)
library(stringr)
library(gt)
library(broom)
library(knitr)
library(MatchIt)
library(arrow)
```

@@

Since the economic reform in China in 1978, China’s GDP per capita has increased more than 29 times [@NBSC2020]. However, this period of rapid economic growth has been accompanied by greater socioeconomic inequality. The increased stratification, along with a number of China specific policies, has pushed the GINI coefficient close to 0.5, and various studies have demonstrated that the majority of China’s wealth is concentrated in the hands of a few [@XieJin2015].

In examining the factors contributing to China’s increasing inequality, a multitude of social scientists highlight China’s education system. The correlation between socioeconomic status and education has been widely scrutinized. According to Esping-Andersen, "in modern capitalist society, the two main exclusionary devices through which the bourgeoisie structures and perpetuates itself as a class, are first, those surrounding the institutions of property, and second, academic or professional qualifications and credentials." [@EspingAndersen1982]

In China, educational attainment directly corresponds to future earnings. Research by Byron and Manaloto showed an approximate 4 percent increase in earnings for each year of schooling [@Byron1990], and Wu and Xie identified a rate of 4.7 percent [@Wu2003]. Jansen and Wu documented a significant rising trend of returns on education during China’s economic reform period from 1978 to 2005 [@Jansen2012]. Concurrently, China significantly expanded its higher education sector, raising the gross enrollment rate from 9.8 percent in 1998 to 24.2 percent in 2009 [@Wang2011]. As of 2022, the enrollment rate stands at 59.7 percent, indicating that China has fully transitioned from elite to mass higher education. Although increased access to higher education and higher returns on education are generally believed to foster social mobility, China’s inequality has worsened. To elucidate this paradoxical outcome, researchers found that the new educational opportunities are predominantly accessed by privileged groups, such as individuals with higher social status, males, and urban populations [@Ou2019].

Previous research has also investigated the reasons behind the educational gap between high and low-income individuals. In the 20th century, many theorists proposed theories examining the link between socioeconomic status and educational achievement. Some suggest that income-based segregation places low-income children in overcrowded classrooms with less adequately trained educators. Others argue that teachers may harbor negative attitudes towards low-income children, anticipating they will perform poorly academically, thereby creating a self-fulfilling prophecy [@Bond1981]. More recently, Duncan and Murnane suggested that the disparity in child enrichment expenditure influences academic performance, giving children from high-income families an advantage [@Duncan2014].

However, while the relationship between financial status and educational attainment is relatively well established, there is a dearth of research, especially in China, to explore the link between family financial status and the rate of return from education. This is an essential connection, as it signifies how effectively education translates into its ultimate outcome: income. The aim of this research is to fill this knowledge gap by examining whether individuals from different socioeconomic backgrounds in the Chinese population receive the same level of returns from education.

The estimand of this investigation is the disparities in educational returns among individuals from different family wealth quartiles during the decade 2020 to 2020, measured in the percent income increase per year of education received.

To arrive at a conclusion, national family and individual data from China Family Panel Studies [@cfps] from the year 2010 to 2020 were extracted. For the data in each CFPS iteration, Propensity Score Matching (PSM) was utilized to balance the control variables across different family wealth quartiles, and separate polynomial regressions were fit to identify the disparities of educational returns across quartiles. It was found that throughout the decade, individuals from wealthier families possessed a significant advantage over individuals from less wealthy families, with disparities reaching more than 3% income increase per year of education between some quartiles.

The body of the paper contains the data, model, results, and discussion sections. In order, @sec-data contains an overview of the data, the variables used, and the programming tools used in my analysis and interpretation of the data. @sec-model describes the model in detail and provides justification for it. @sec-results contains a run-down and explanation of the results and findings from the model, and @sec-discussion contains a discussion of the implications and limitations of the result, among others. The appendices contain a detailed examination of the methodologies of CFPS, as well as aspects of the data and analysis excluded from the body.

@@

I used the statistical programming language R [@R] for the graphing, analysis and presentation of the project as a whole. Caret [@caret], glmnet [@glmnet; @glmnet2; @glmnet3], nnet [@nnet], tidyverse [@tidyverse], car [@car] MatchIt [@matchit] and arrow [@arrow] were used directly in the analysis of the data. Stringr [@stringr], dplyr [@dplyr], gt [@gt], knitr [@knitr], and styler [@styler] were used in the presentation and/or styling of the data, graphs, and paper as a whole.

All data utilized in this study were sourced from various iterations of the China Family 
Community Survey [@cfps]. The CFPS is a nationwide community survey funded by 985 Program of Peking University and carried out by the Institute of Social Science Survey of Peking University. CFPS encompasses approximately 15,000 families across 25 provinces and aims to tracks them permanently since 2010. Its main objective is to gather data across three dimensions: individuals, households, and communities. To achieve this, CFPS comprises four sub-surveys that thoroughly investigates families, adults, and children, gathering data for thousands of variables. Given its expansive and representative sample and a wide range of variables, CFPS has become a staple dataset for research in Chinese social sciences. Alternative data sources for this study, such as China General Social Survey (CGSS) and China Social Survey (CSS) are also reliable but do not provide the same width of variables, especially concerning family units. For the purposes of this research, data was extracted from CFPS for the years 2010, 2012, 2014, 2016, 2018, and 2020.

@@

@@

The variables presented in this section are variables in the polynomial regression model to identify returns to education, which is a coefficient of the model. This will be specified in @sec-model.

The outcome variable of the model is the natural logarithm of individuals' hourly wages. @fig-outcome shows its distribution in the aggregated survey data from all CFPS iterations between 2010 and 2020.

@@

data_2020 <- read_parquet(here::here("data/02-analysis_data/2020_analysis_data.parquet"))
data_2018 <- read_parquet(here::here("data/02-analysis_data/2018_analysis_data.parquet"))
data_2016 <- read_parquet(here::here("data/02-analysis_data/2016_analysis_data.parquet"))
data_2014 <- read_parquet(here::here("data/02-analysis_data/2014_analysis_data.parquet"))
data_2012 <- read_parquet(here::here("data/02-analysis_data/2012_analysis_data.parquet"))
data_2010 <- read_parquet(here::here("data/02-analysis_data/2010_analysis_data.parquet"))

# Make a combined dataset that includes all survey iterations
all_combined <- bind_rows(
  data_2020, data_2018, data_2016, data_2014, data_2012, data_2010
)
```

@@

# Outcome variable: ln hourly wage
ggplot(all_combined, aes(x = ln_hourly_wage)) +
  geom_histogram(
    fill = "#2c7bb6", # Aesthetic fill color (steel blue)
    color = "white", # Border color for bins
    alpha = 1 # Transparency of the fill
  ) +
  labs(
    x = "ln Hourly Wage",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  ) +
  # Add a vertical line for the mean percentage support
  geom_vline(
    aes(xintercept = mean(ln_hourly_wage, na.rm = TRUE)),
    color = "#d73027", # Aesthetic color (red)
    linetype = "dashed",
    size = 1
  ) +
  # Annotate the mean value on the plot
  annotate(
    "text",
    x = mean(all_combined$ln_hourly_wage, na.rm = TRUE) - 7, # Position text slightly to the right of the mean line
    y = Inf, # Position text at the top of the plot
    label = paste("Mean:", round(mean(all_combined$ln_hourly_wage, na.rm = TRUE), 1)),
    vjust = 2, # Vertical adjustment
    color = "#d73027",
    size = 5,
    fontface = "bold"
  )
```

@@

### End Date

```{r}
#| label: fig-dates-past
#| fig-cap: Distribution of Poll End Date, 2020 Cycle
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined %>% filter(cycle == 2020), aes(x = end_date)) +
  geom_histogram(binwidth = 7, fill = "#1f78b4", color = "white", alpha = 0.7) +
  labs(
    x = "End Date",
    y = "Number of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{=tex}
\vspace{1em}
\FloatBarrier
```
```{r}
#| label: fig-dates-current
#| fig-cap: Distribution of Poll End Date, 2024 Cycle
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined %>% filter(cycle == 2024), aes(x = end_date)) +
  geom_histogram(binwidth = 7, fill = "#1f78b4", color = "white", alpha = 0.7) +
  labs(
    x = "End Date",
    y = "Number of Polls"
  ) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

The end date indicates the date that a poll ends on. The dates are graphed separately between the 2020 and 2024 cycles for ease of viewing. Notice how there is a disproportionately large uptick in the number of polls that end on the end of a month.

@@

Sample size, finally, is a measure of the number of participants in each poll. Around 1000 participants seems to overwhelmingly be the final survey size that pollsters decide on, with only single-digit numbers of polls (with rating \>= 2.5) boasting more than four thousand respondents.

@@

### Logistic models

Define $y_i$ as the binary outcome variable indicating whether Harris wins (1) or not (0) for the $i^{th}$ observation. $x_{1i}$ is the end_date for the $i^{th}$ observation. $x_{2i}$ is the binary value of has_sponsor for the $i^{th}$ observation. $x_{3i}$ is the transparency_score for the $i^{th}$ observation. $x_{4i}$ is the sample_size for the $i^{th}$ observation.

@@

```{=tex}
\begin{align} y_i &\sim \text{Bernoulli}(p_i) 
\\ \text{logit}(p_i) &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i} \end{align}
```
Line (1) specifies that the outcome variable $y_i$ follows a Bernoulli distribution with probability $p_i$ of success (i.e., Harris winning). Line (2) links the linear combination of predictors to the probability $p_i$, where $\beta_0$ is the intercept term, and $\beta_1, \beta_2, \beta_3, \beta_4$ are coefficients corresponding to each predictor variable above.

@@

### Multiple linear regression models

Define $y_i$ as the percentage national support for Harris or Trump for the $i^{th}$ observation. $x_{1i}$ is the end_date for the $i^{th}$ observation. $x_{2i}$ is the binary value of has_sponsor for the $i^{th}$ observation. $x_{3i}$ is the transparency_score for the $i^{th}$ observation. $x_{4i}$ is the sample_size for the $i^{th}$ observation.

@@

```{=tex}
\begin{align} y_i &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i} \end{align}
```
Where $\beta_0$ is the intercept term, and $\beta_1, \beta_2, \beta_3, \beta_4$ are coefficients corresponding to each predictor variable.

@@

roc_obj_adj <- roc(
  combined_national_adj$harris_win,
  combined_national_adj$harris_win_prob
)
auc_value_adj <- auc(roc_obj_adj)
```

For the analysis of the national popular vote, the logistic model of adjusted polls possess an AUC value of 0.98, and the logistic model of unadjusted polls have an AUC of 0.94, indicating high reliability of categorization. Their respective ROC plots are presented as @fig-log and @fig-logadj.

@@

mean_probabilities %>%
  gt() %>%
  fmt_number(
    columns = vars(Harris_Win_Probability),
    decimals = 2
  ) %>%
  cols_label(
    State = "State",
    Harris_Win_Probability = "Harris Win Probability (%)",
    electoral_votes = "Electoral Votes"
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 12,
    table.border.top.width = px(2),
    table.border.bottom.width = px(2)
  )
```

Combining with the fact that Harris requires 44 electoral votes to win and Trump requires 51, we have found using the probability function that Harris has a 38.3 chance to acquire more than 44 electoral votes, according to the probabilities and electoral votes in @tbl-result. On the other hand, Trump has a 61.7 probability to acquire more than 51 votes. Thus, our conclusion is that while both candidates possess a significant chance of winning, Trump has the greater probability of acquiring the necessary electoral votes from the seven swing states on election day.

@@

# create the plot that contains scattered points of adjusted percentages
# and lines for fitted values
ggplot(combined_data_adj, aes(x = end_date, y = pct_adj, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

@@

```{r}
#| label: fig-lr
#| fig-cap: Unadjusted Popular Vote Prediction, 2024
#| echo: false
#| message: false
#| warning: false
# Unadjusted support percentages
ggplot(combined_data, aes(x = end_date, y = pct, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

@@

# Discussion {#sec-discussion}

@@

# Appendix {.unnumbered}

## Pollster methodology overview and evaluation

The Quinnipiac University Poll conducts independent polling in swing states. To analyze their methodology, we look specifically into their October 2024 Pennsylvania polls [@quinnipiac].

@@

## Idealised methodology

With a \$100,000 budget, our approach focuses on producing accurate state-level estimates in key battleground states, which would then inform our national forecast. We prioritize Pennsylvania, Michigan, Wisconsin, Georgia, Arizona, and Nevada, allocating resources proportionally based on each state's electoral importance and expected margin of victory.

Our sampling strategy employs both probability and non-probability methods. Probability sampling (60% of budget, i.e. \$60k) means every member of our target population has a known, non-zero chance of being selected - this allows us to calculate proper margins of error and make statistical inferences about the population. For this, we use dual-frame random digit dialing (RDD) for phone surveys and address-based sampling (ABS) for mail-to-web recruitment. RDD involves generating random phone numbers within active area codes, while ABS uses the U.S. Postal Service's delivery database as a sampling frame, both proven methods of sampling [@cummings]. The ABS approach helps reach households without reliable phone access. We stratify our sample by geography, demographics, and previous voting patterns to ensure representation across key subgroups - meaning we divide the population into these subgroups and sample from each independently.

For non-probability sampling (40% of budget, i.e. \$40k), where respondents' selection probabilities are unknown and not everyone has a chance of being selected, we recruit through multiple online panel vendors and use targeted social media advertising to reach traditionally underrepresented groups. While this approach introduces potential selection bias because participants self-select into the sample, it helps reach younger voters who are less responsive to traditional survey methods [@nonprobabilitysampling]. We implement quota sampling within these non-probability samples to match key demographic targets - for example, stopping collection from certain demographic groups once their quota is filled.

Respondent recruitment uses multiple contact methods - mail, email, text, and phone - with attempts made at different times and days to maximize response rates. We offer a \$10 gift card incentive for completed surveys and provide both English and Spanish language options. This mixed-mode contact strategy helps reduce non-response bias by providing multiple ways to participate.
